\pkg{mlr3spatiotempcv}: Spatiotemporal Resampling Methods for \pkg{mlr3}

### 2021-04-04 paper/paper.Rmd
refinements

```{r setup, echo=FALSE}
# important for correct figure placement
# see https://stackoverflow.com/a/49395389/4185785
knitr::knit_hooks$set(plot = function(x, options) {
  knitr::hook_plot_tex(x, options)
})
# knitr::opts_chunk$set(cache = FALSE)
```

@@

@@

Every partitioning method in [mlr3spatiotempcv](https://mlr3spatiotempcv.mlr-org.com) comes with a generic `plot()` method to visualize the created groups.
In a 2D space this happens via [ggplot2](https://ggplot2.tidyverse.org) [@ggplot2] while for spatiotemporal methods 3D visualizations are created via [plotly](https://github.com/ropensci/plotly) [@plotly].

Visualization of partitioning methods is a vital part of \pkg{mlr3spatiotempcv} as most upstream packages lack support for such.
The semantics are presented first because plotting will be used extensively when presenting the integrated methods.

Unless specified by the user, the coordinate reference system (CRS) defaults to EPSG code 32717 (WGS84).
This is because a lat/lon based CRS is better suited for plotting purposes than a Mercator (UTM) one due tho a more descriptive axes labeling.
Setting an appropriate CRS for the given data *during task construction* is very important.
Even though EPSG 32717 is a good fallback and often used for visualization purposes, spatial offsets of up to multiple meters may occur if the wrong CRS was passed initially.

The following example creates a spatiotemporal task from scratch.
The data used here is also available as a built-in task via `tsk("ecuador")`.

Because the spatial grouping of the k-means based approach (`"spcv_coords"`) (Fig. \@ref(fig:vis-spcv)) contrasts visually very well to the NSpCV (random) partitioning (Fig. \@ref(fig:vis-nspcv)), this method was chosen to showcase the 2D plotting capabilities of \pkg{mlr3spatiotempcv}.

The partitions shown in the following are also used in the case study in section \@ref(sec:case-study).

@@

At the time of writing, the following spatiotemporal resampling methods (Tab. \@ref(tab:sptcv-methods)) were available across different R packages.
This list might increase over time and it is recommended to consult the homepage of \pkg{mlr3spatiotempcv} for an up-to-date list.
The internal identifier for the respective method in \pkg{mlr3spatiotempcv} is shown in column `Notation`.

@@

autoplot(resampling_buffer,
  size = 0.8, task = task, fold_id = 1, crs = 32717) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

autoplot(resampling_block_random,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

autoplot(resampling_block_systematic,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

autoplot(resampling_block_checkerboard,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

autoplot(resampling_block_rows_cols,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

autoplot(resampling_env,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

autoplot(resampling_env_multi,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

autoplot(resampling_coords,
  size = 0.8, fold_id = 1, task = task, crs = 32717) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

autoplot(resampling_cstf_time,
  fold_id = 5, task = task_spt, crs = 32717, plot3D = TRUE
)
```

@@

autoplot(resampling_cstf_loc,
  fold_id = 5, task = task_spt, crs = 32717, plot3D = TRUE)
```

@@

autoplot(resampling_cstf_time_loc,
  fold_id = 4, task = task_spt, crs = 32717, plot3D = TRUE,
  show_omitted = TRUE)
```

@@

<!-- env block: hard to decide which variable to use for clustering -->
<!-- env block: more than one variable may lead to unforeseeable results -->
Method `"spcv_env"` is tricky in that it requires the user to limit the clustering to (a) specific variable(s).
This implies that only this/these variable(s) suffer from autocorrelation and others can be ignored.
However, the influence of a single variable is hard to estimate and the same applies for estimating the influence of the non-accounted variables which might add substantial bias to the CV.

@@

>>>>

```{r setup, echo=FALSE, purl=FALSE}
# important for correct figure placement
# see https://stackoverflow.com/a/49395389/4185785
knitr::knit_hooks$set(plot = function(x, options) {
  knitr::hook_plot_tex(x, options)
})
# knitr::opts_chunk$set(cache = FALSE)
```

```{r setup-2, echo=FALSE, purl=TRUE}
# for knitr spint later
knitr::opts_chunk$set(fig.path = "./paper/figs/")
```

@@

In addition, \pkg{mlr3spatiotempcv} adds 2D and 3D visualization possibilities for all implemented methods.
The authors believe that the possibility for creating graphical representations of partitions is important, especially if spatio-temporal clustering or other grouping types were used to create the groups.

@@

Every partitioning method in [mlr3spatiotempcv](https://mlr3spatiotempcv.mlr-org.com) comes with generic `plot()` and `autoplot()` (\pkg{ggplot2} equivalent) method to visualize the created groups.
In a 2D space this happens via package [ggplot2](https://ggplot2.tidyverse.org) [@ggplot2] while for spatiotemporal methods 3D visualizations are created via package [plotly](https://github.com/ropensci/plotly) [@plotly].

Visualization of partitioning methods is a vital part of \pkg{mlr3spatiotempcv} as most upstream packages lack support for such.
Hence, plotting semantics are presented first because plotting will be used extensively when presenting the integrated methods.

Unless specified by the user, the coordinate reference system (CRS) defaults to EPSG code 4326 (WGS84).
This is because a lat/lon based CRS is better suited for plotting purposes than a Mercator (UTM) one due tho a more descriptive axes labeling.
Setting an appropriate CRS for the given data *during task construction* is very important.
Even though EPSG 4326 is a good fallback and often used for visualization purposes, spatial offsets of up to multiple meters may occur if the wrong CRS was passed initially.

The following example creates a spatiotemporal task from scratch.
This task is also available as a built-in task via `tsk("ecuador")`.

Because the spatial grouping of the k-means based approach (`"spcv_coords"`) (Fig. \@ref(fig:vis-spcv)) contrasts visually very well to NSpCV (random) partitioning (Fig. \@ref(fig:vis-nspcv)), this method was chosen to showcase the 2D plotting capabilities of \pkg{mlr3spatiotempcv}.

The partition plots shown in the following are also used in the case study in section \@ref(sec:case-study).

@@

At the time of writing, the following spatiotemporal resampling methods (Tab. \@ref(tab:sptcv-methods)) were available across different R packages.
This list might increase over time and it is recommended to check the homepage of \pkg{mlr3spatiotempcv} for an up-to-date list.
In Tab. \@ref(tab:sptcv-methods), the internal identifier for the respective method in \pkg{mlr3spatiotempcv} is shown in column "Notation".
Some of the methods listed are able to produce different partitioning results by setting different options, hence the actual number of available methods is higher than the one shown in Tab. \@ref(tab:sptcv-methods).

@@

plot(resampling_buffer,
  size = 0.8, task = task, fold_id = 1, crs = 32717) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

plot(resampling_block_random,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

plot(resampling_block_systematic,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

plot(resampling_block_checkerboard,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

plot(resampling_block_rows_cols,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

plot(resampling_env,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

plot(resampling_env_multi,
  size = 0.8, fold_id = 1, task = task, crs = 32717,
  show_blocks = TRUE, show_labels = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

plot(resampling_coords,
  size = 0.8, fold_id = 1, task = task, crs = 32717) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

plot(resampling_cstf_time,
  fold_id = 5, task = task_spt, crs = 32717, plot3D = TRUE
)
```

@@

plot(resampling_cstf_loc,
  fold_id = 5, task = task_spt, crs = 32717, plot3D = TRUE)
```

@@

plot(resampling_cstf_time_loc,
  fold_id = 4, task = task_spt, crs = 32717, plot3D = TRUE,
  show_omitted = TRUE)
```

@@

<!-- env block: hard to decide which variable to use for clustering -->
<!-- env block: more than one variable may lead to unforeseeable results -->
Method `"spcv_env"` is complicated in that it requires the user to limit the clustering to (a) specific variable(s).
This implies that only this/these variable(s) suffer from autocorrelation and others can be ignored.
However, the influence of a single variable is hard to estimate and the same applies for estimating the influence of the non-accounted variables which might add substantial bias to the CV.

@@

```{r echo=FALSE}
sessionInfo()
```
